Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-10-20 manuscript.Rmd
Update manuscript.Rmd
Apply andreas' suggestions
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

We outline two interpretations of preregistration.
The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which otherwise may be reduced by researchers' degrees of freedom, p-hacking, and suchlike.
The second interpretation is our main contribution:
We argue that contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk.
This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
In reality, surely researchers differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
The goal is therefore to reason about the space of possible measures researchers might apply.
However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence confirmation, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is equally reasonable; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
We call this probability of detecting evidence, given that the hypothesis holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
However, we latter discuss how issues detectability must be considered for the issue of preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, decreasing it can increase the posterior probability.
In other words, high risk, high reward.

@@

Let us note some interesting properties of theoretical risk $P(\neg{}E|\neg{}H)$.
First, increasing theoretical risk leads to higher posterior probability $P(H|E)$, our objective.
Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (an inversion of statistical relevancy).
Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
In other words, observing a sure fact does not lend credence to a hypothesis.

>>>>

Now, we outline two possible perspectives on the utility of preregistration.
The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which otherwise may be reduced by researchers' degrees of freedom, p-hacking, and suchlike.
The second interpretation is our main contribution:
We argue that contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk.
This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called _confirmation as firmness_; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
In reality, researchers surely differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
The goal is therefore to reason about the space of possible measures researchers might apply.
However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence confirmation, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is equally reasonable; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
We call this probability of detecting evidence, given that the hypothesis holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
We later discuss how issues of detectability must be considered in a preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, decreasing it can increase the posterior probability.
In other words, high risk, high reward.

@@

Let us note some interesting properties of theoretical risk $P(\neg{}E|\neg{}H)$.
First, increasing theoretical risk leads to higher posterior probability $P(H|E)$, our objective.
Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (the inverse of statistical relevancy).
Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
In other words, observing a sure fact does not lend credence to a hypothesis.
